Initial Analysis

### 2019-09-19 Initial_Analysis.Rmd
Finished function to estimate the species status according to the SAU method

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "data.table", # Read dataframe (Fast!)
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  "wesanderson",
  "circlize", # For circular plot
  "ggimage", #for reading images to the circular plot
  # "cowplot",
  # "rgdal",
  # "RColorBrewer",
  # "knitr",
  # "kableExtra",
  "ggrepel", # for nice plot labels
  "ggsflabel", # for nice sf_plots labels
  # "gridExtra",
  # "ggmap",
  # "rgeos",
  # "stringr",
  # "stringer",
  "spdep", # for poly2nb
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  "purrr",#Spatial analysis
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  # "png", # For reading plots in chunk codes
  # "grid" # For reading plots in chunk codes
  "parallel", # for parallelization
  "taxize", # For getting species names
  "rfishbase" # for species ecosystem affinity
)

@@

#### For testing 
# Spp <- 600004  # Anchovie Equator, Chile and Peru
# Spp <- 600107 # Tuna Many RFMOs
# Spp <- 600464 # Discrete species 
# Spp <- 600051 # One of Pierrs massive (wrong) discrete spp
# x <- c(600464, # Discreat spp in one nation
      # 600059, # nowhere to be found spp
       # 601352, # Discrete spp in two nations
       # 600004) # transboundary spp
# Model <- "All"
# Coord <- CoorG
# 
# # For EEZ
# Neighbours <- Neighbours_Data
# Index_Code <- Index_Code
#  
# __________________________________
# __________________________________

@@

  }
  ### Function return
  EstTransIndex=Indexes # A dataframe with

@@

@@

@@

World_sf_si %>% 
  filter(Territory == "Côte d'Ivoire")

@@

World_sf_si_land <- World_land_sf %>%
  st_transform(crs = 4326) %>% # 4326
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>%  # Include transboundary species
  left_join(Catch_Country, by ="Territory") # Include total specices catch

@@

World_sf_si$Bins <- cut(World_sf_si$n_trans_spp,breaks = 4)
World_sf_si_land$Bins <- cut(World_sf_si_land$Total_Catch,breaks = 4)

@@

Plot <- ggplot() +
  geom_sf(data = World_sf_si_land,aes(fill = Bins), colour = "grey80") +
  geom_sf(data = World_sf_si,aes(fill = Bins),colour = "grey20") + # n_trans_spp or Bins
  geom_sf(data = Grey_land,aes(fill = Bins),fill = "grey80") + # n_trans_spp or Bins
  # scale_fill_gradientn(colours = pal,
  #                      na.value = "white") +
  # scale_fill_manual(values = c(rev(wes_palette(n = 4, name = "Darjeeling1")),
  #                              rev(wes_palette(n = 4, name = "Darjeeling2"))
  #                              ),
  #                    na.value = "white"
  # ) +
  scale_fill_manual("",values = c("#5BBCD6","#5BBCD6","#00A08A","#00A08A","#F2AD00","#F2AD00","#FF0000","#FF0000"),
  na.value = "white"
  ) +
  # scale_fill_manual(values = rev(wes_palette(n = 5, name = "Darjeeling1")),
  #                    na.value = "white"
  # ) +
  ggtheme_map()

Plot_Name <- paste(Jurisdiction,"_100_25_Bins_Landb.png", sep = "")

@@

Countries_Data_Plot <- Tresholded_Data %>% 
  group_by(Territory = CT_fishing_entity) %>% 
  summarise(value = length(unique(TaxonKey))) %>% 
  left_join(UN_Regions, by ="Territory") %>%
  mutate(group = ifelse(is.na(georegion),"Other",paste(georegion))) %>% 
  rename(FishingRegion=Territory) %>% 
  left_join(Clean_sau_eez_fes) %>% 
  group_by(individual = fishing_entity,group) %>% 
  summarise(Total_trans = sum(value,na.rm=T)) %>% 
  mutate(value = ifelse(Total_trans > 100,100,Total_trans)) %>% 
  filter(!is.na(individual)) %>% 
  select(individual,group,value) %>% 
  arrange(group,value)

@@

>>>>

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "data.table", # Read dataframe (Fast!)
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  "wesanderson",
  "circlize", # For circular plot
  "ggimage", #for reading images to the circular plot
  # "cowplot",
  # "rgdal",
  # "RColorBrewer",
  # "knitr",
  # "kableExtra",
  "ggrepel", # for nice plot labels
  "ggsflabel", # for nice sf_plots labels
  # "gridExtra",
  # "ggmap",
  # "rgeos",
  # "stringr",
  # "stringer",
  "spdep", # for poly2nb
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  "purrr",#Spatial analysis
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  # "png", # For reading plots in chunk codes
  # "grid" # For reading plots in chunk codes
  "parallel", # for parallelization
  "taxize", # For getting species names
  "rfishbase", # for species ecosystem affinity
  "zoo" #for runing mean
)

@@

#### For testing 
# Spp <- 600004  # Anchovie Equator, Chile and Peru
# Spp <- 600107 # Tuna Many RFMOs
# Spp <- 600464 # Discrete species 
# Spp <- 600051 # One of Pierrs massive (wrong) discrete spp
# x <- c(600464, # Discreat spp in one nation
      # 600059, # nowhere to be found spp
       # 601352, # Discrete spp in two nations
       # 600004) # transboundary spp
# Model <- "All"
# Coord <- CoorG
# 
# # For EEZ
# Neighbours <- Neighbours_Data
# Index_Code <- Index_Code

# __________________________________ #

@@

  if(nrow(Indexes) > 0 & Save == "N"){
  EstTransIndex=Indexes # A dataframe with   
  }

  ### Function return
  EstTransIndex=Indexes # A dataframe with

@@

## Function `EstSppStatus`

```{r Species_Status_fun, eval = T, echo =F}

EstSppStatus <- function(Spp,Coord){

  Data <- GetSppDist(Spp, Model = "SAU_C", Coord = Coord) %>% 
    gather("Year","Value",CATCH.1:CATCH.65) %>% 
    arrange(INDEX) %>% 
    mutate(Year = rep(seq(1951,2014,1),263250)) %>% 
    left_join(Index_Code,
              by = "INDEX") %>% 
    filter(!is.na(EEZID),
           Value > 0) %>% 
    group_by(Species,Year,Territory) %>% # Total catch per country
    summarise(Total_Catch = sum(Value,na.rm=T)) %>% 
    group_by(Species,Territory,Year) %>% 
    mutate(RMean = rollmean(x = Total_Catch, 3,
                            align = "right",
                            fill = Total_Catch)
           ) %>% # Running mean
    group_by(Species,Territory) %>% 
    select(-Total_Catch)

  # Tresholds to evaluate stocks if these are not met, then soecies status = "No Status"
  Tresholds_Data <- Data %>% 
    arrange(Territory) %>% 
    group_by(Species,Territory, grp = cumsum(c(1, diff(Year) != 65))) %>% 
    group_by(Species,Territory) %>% 
    summarise(Years_appart = ifelse(max(Year)-min(Year) >= 0,"Yes","No"), # Step 1. first and last reported landings are at least 10 years apart
              Consecutive = ifelse(max(grp) >= 5,"Yes","No"), # Step 2. There are at least 5 consecutive catch years
              Catch_over = ifelse(max(RMean >= 1000),"Yes","No") # Step 3. 3) The catch in a particular area (LME) is at least 1,000 tonnes.
    ) %>% 
    gather("Step","Treshold",Years_appart:Catch_over) %>% 
    group_by(Species,Territory) %>% 
    summarise(Met_Criteria = paste(unique(Treshold),collapse =";")
              )

  # Remove species that do not met the criterias

    No_Status_Data <- Tresholds_Data %>% 
      filter(Met_Criteria != "Yes") %>% 
      select(Species,Territory) %>% 
      mutate(Status = "No Status")

  # Evaluate species that met the criteria

    # Get Variables to perform status metric
    Status_Data <- Data %>% 
      anti_join(No_Status_Data, # Removes no status data
                by = c("Species","Territory")
      ) %>% 
      # Get Max catch and max catch year
      group_by(Species,Territory) %>% 
      top_n(1,RMean) %>% 
      rename(Max_Peak = RMean,
             Y_Max_Peak = Year) %>% 
      # Get Min catch after peak catch
      left_join(Data,
                by = c("Species","Territory")
                ) %>% 
      filter(Year > Y_Max_Peak) %>% 
      group_by(Species,Territory) %>% 
      top_n(-1,RMean) %>% 
      rename(Post_Max_Min = RMean,
             Y_Post_Max_Min = Year) %>% 
      left_join(Data,
                by = c("Species","Territory")
                ) %>% 
      # Final Evaluation 
      mutate(Status = ifelse(Year > Y_Post_Max_Min & Post_Max_Min < (Max_Peak*0.10) & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50)),"Rebuilding",
                             ifelse((Year < Y_Max_Peak & RMean <= (Max_Peak*0.50) | Y_Max_Peak == max(Year)),"Developing",
                                    ifelse(RMean > Max_Peak*0.50,"Max Exploited",
                                           ifelse(Year > Y_Max_Peak & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50)),"Over Exploited",
                                                  ifelse(Year > Y_Max_Peak & RMean < Max_Peak*0.10,"Collapsed",
                                                         "No Status")
                                           )
                                    )
                             )
      )
      ) %>% 
      filter(Year > 2004) %>% # last 10 years of data
      group_by(Species,Territory,Status) %>% 
      summarise(n_Status= n()) %>% 
      group_by(Species,Territory) %>% 
      top_n(1,n_Status) %>%
      select(-n_Status)

    # If both datasets have value RETURN BOTH
    if(nrow(No_Status_Data) > 0 & nrow(Status_Data) > 0){

      Fun_Result <- bind_rows(No_Status_Data,
                              Status_Data)
    }

    # If only No status data, then just return that
    if(nrow(No_Status_Data) > 0 & nrow(Status_Data) == 0){
      return(No_Status_Data)
    }else{
      return(Status_Data)
    }

# Testing
# Test_1 <- EstSppStatus(Species_List[1],CoorG)

Species_Status <- bind_rows(
    mclapply(
      Species_List,
      FUN = EstSppStatus,
      Coord = CoorG  # Gabriels coordinate system for read species dist
    )
  )

@@

  # Read Sau Value Data (Thank you Tim!)
  Sau_Revenue_Data <- fread(paste(Data_Path,"Distribution_Data/sau_catch_value_country_taxon_JEPA.csv", sep ="")) %>% 
    group_by(fishing_entity_id,TaxonKey=taxon_key) %>% 
    summarise(Mean_Value = mean(value,na.rm=T))

@@

#  Top 10 Larger EEZs...
# World_sf_si %>% 
#   arrange(desc(Area_km2)) %>% 
#   head(10)

@@

# Or revenue data %>%

Tresholded_T_Data_x <- Tresholded_T_Data %>% 
  select(TaxonKey, Territory=CT_fishing_entity)

Revenue_Country <- sau_eez_fes %>% 
  group_by(fishing_entity_id, Territory = fishing_entity) %>% 
  summarise(n=n()) %>% 
  select(-n) %>% 
  left_join(Tresholded_T_Data_x) %>% 
  left_join(Sau_Revenue_Data) %>% 
  group_by(Territory) %>% 
  summarise(Total_USD = sum(Mean_Value,na.rm=T)) %>% 
  mutate(Territory = ifelse(Territory == "USA","United States of America",Territory))

hist(Revenue_Country$Total_USD)

World_sf_si_land <- World_land_sf %>%
  st_transform(crs = 4326) %>% # 4326
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>%  # Include transboundary species
  left_join(Revenue_Country, by ="Territory") # Include total specices catch

@@

World_sf_si$Bins <- cut(World_sf_si$n_trans_spp,breaks = 4)
World_sf_si_land$Bins <- cut(World_sf_si_land$Total_USD,breaks = 4)

World_sf_si_land <- World_sf_si_land %>% 
  arrange(Total_USD)

World_sf_si <- World_sf_si %>% 
  arrange(n_trans_spp)

@@

Plot <- ggplot() +
  geom_sf(data = World_sf_si_land,aes(fill = Bins), colour = "grey80",size = 0.1) +
  geom_sf(data = World_sf_si,aes(fill = Bins),colour = "grey20", size = 0.1) + # n_trans_spp or Bins
  geom_sf(data = Grey_land,aes(fill = Bins),fill = "grey80", colour = "grey50",size = 0.1) + # n_trans_spp or Bins
  # scale_fill_gradientn(colours = pal,
  #                      na.value = "white") +
  # scale_fill_manual(values = c(rev(wes_palette(n = 4, name = "Darjeeling1")),
                               # rev(wes_palette(n = 4, name = "Darjeeling2"))
                               # ),
                     # na.value = "white"
  # ) +
  scale_fill_manual("",values = c("#5BBCD6","#5BBCD6","#00A08A","#00A08A","#F2AD00","#F2AD00","#FF0000","#FF0000"),
  na.value = "white"
  ) +
  # scale_fill_manual(values = rev(wes_palette(n = 5, name = "Darjeeling1")),
  #                    na.value = "white"
  # ) +
  ggtheme_map()

Plot_Name <- paste(Jurisdiction,"_100_25_Bins_Value.png", sep = "")

@@

Countries_Data_Plot <- Tresholded_T_Data %>% 
  group_by(Territory = CT_fishing_entity) %>% 
  summarise(value = length(unique(TaxonKey))) %>% 
  left_join(UN_Regions, by ="Territory") %>%
  mutate(group = ifelse(is.na(georegion),"Other",paste(georegion))) %>% 
  rename(FishingRegion=Territory) %>% 
  left_join(Clean_sau_eez_fes) %>% 
  group_by(individual = fishing_entity,group) %>% 
  summarise(Total_trans = sum(value,na.rm=T)) %>% 
  mutate(value = ifelse(Total_trans > 100,100,Total_trans)) %>% 
  filter(!is.na(individual)) %>% 
  select(individual,group,value) %>% 
  arrange(group,value)

@@

### Species status (SAU methods)

Note that (n), the number of ‘stocks’ is defined as a time series of a given species, genus or family (higher and pooled groups have been excluded) for which the first and last reported landings are at least 10 years apart, for which there are at least 5 years of consecutive catches and for which the catch in a given area is at least 1000 tonnes.

 1) The first and last reported landings are at least ten years apart;

2) There are at least five years of consecutive catches; and

3) The catch in a particular area (LME) is at least 1,000 tonnes.

Once those tresholds are met, the status of the species is determined as follows:

- Rebuilding (Recovering)	Year of landing > year of post-max. min. landing AND post-max. min. landing < 10% of max. landing AND landing is 10-50% of max. landing
- Developing	Year of landing < year of max. landing AND landing is < or = 50% of max. landing OR year of max. landing = final year of landing
- Exploited	Landing > 50% of max. landing
- Over exploited	Year of landing > year of max. landing AND landing is between 10-50% of max. landing
- Collapsed	Year of landing > year of max. landing AND landing is < 10% of max. landing

*Notes* = 3 years average mean

```{r Species_Status, eval = T, echo =F}

Spp <- GetSppDist(600004, Model = "SAU_C", Coord = CoorG)

Data <- Spp %>% 
  gather("Year","Value",CATCH.1:CATCH.65) %>% 
  arrange(INDEX) %>% 
  mutate(Year = rep(seq(1951,2014,1),263250)) %>% 
  left_join(Index_Code) %>% 
  filter(!is.na(EEZID),
         Value > 0) %>% 
  group_by(Species,Year,EEZID,Territory) %>% # Total catch per country
  summarise(Total_Catch = sum(Value,na.rm=T)) %>% 
  group_by(Species,Territory,EEZID,Year) %>% 
  mutate(RMean = rollmean(x = Total_Catch, 5, align = "right", fill = Total_Catch)) %>% 
  group_by(Species,Territory,EEZID)

# head(Data)

# Tresholds to evaluate stocks if these are not met, then soecies status = "No Status"
 Tresholds <- Data %>% 
  arrange(Territory) %>% 
  group_by(Species,Territory, grp = cumsum(c(1, diff(Year) != 65))) %>% 
  group_by(Species,Territory) %>% 
  summarise(Years_appart = ifelse(max(Year)-min(Year) >= 0,"Yes","No"), # Step 1. first and last reported landings are at least 10 years apart
            Consecutive = ifelse(max(grp) >= 5,"Yes","No"), # Step 2. There are at least 5 consecutive catch years
            Catch_over = ifelse(max(RMean >= 1000),"Yes","No") # Step 3. 3) The catch in a particular area (LME) is at least 1,000 tonnes.
            ) %>% 
   gather("Step","Treshold",Years_appart:Catch_over) %>% 
   filter(Treshold == "Yes")

 # Evaluate stocks

 # Filter data that does not satisfy tresholds

 Landings_Peaks <- Data %>% 
  group_by(Species,Territory) %>% 
  summarise(Max_Peak = max(RMean,na.rm=T))

 Eval_Data <- Data %>% 
  semi_join(Tresholds,
            by = c("Species","Territory")) %>% 
   left_join(Landings_Peaks,
             by = c("Species","Territory"))

 Peak_Year <- Eval_Data %>% 
   filter(RMean == Max_Peak) %>% 
   select(Species,Territory,EEZID,Y_Max_Peak = Year)

 Post_Max_Min <- Eval_Data %>% 
   left_join(Peak_Year) %>% 
   filter(Year > Y_Max_Peak) %>% 
   group_by(Species,Territory,EEZID) %>% 
   summarise(Post_Max_Min = min(RMean,na.rm=T))

 Y_Post_Max_Min <- Eval_Data %>% 
   left_join(Post_Max_Min) %>% 
   filter(RMean == Post_Max_Min) %>% 
   select(Species,Territory,EEZID,Y_Post_Max_Min = Year)

 Data_Status <- Eval_Data %>% 
   left_join(Peak_Year,
             by = c("Species","Territory","EEZID")) %>%
   left_join(Post_Max_Min,
             by = c("Species","Territory","EEZID")) %>% 
   left_join(Y_Post_Max_Min,
             by = c("Species","Territory","EEZID")) %>% 
   rename(Y_RMean = Year) %>% 
   mutate(Status = ifelse(Y_RMean > Y_Post_Max_Min & Post_Max_Min < (Max_Peak*0.10) & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50)),"Rebuilding",
                          ifelse((Y_RMean < Y_Max_Peak & RMean <= (Max_Peak*0.50) | Y_Max_Peak == max(Y_RMean)),"Developing",
                                 ifelse(RMean > Max_Peak*0.50,"Max Exploited",
                                        ifelse(Y_RMean > Y_Max_Peak & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50)),"Over Exploited",
                                               ifelse(Y_RMean > Y_Max_Peak & RMean < Max_Peak*0.10,"Collapsed",
                                                      "No Status")
                                        )
                                 )
                          )
   )
   ) %>% 
   filter(Y_RMean > 2004) %>% 
   group_by(Species,EEZID,Territory,Status) %>% 
   summarise(n_Status= n()) %>% 
   group_by(Species,EEZID,Territory) %>% 
   top_n(1,n_Status)

 # Manual estimation

 data.frame(
   Y_RMean = 1980,
   RMean = 642187,
   Max_Peak = 1298712,
   Y_Max_Peak = 1990,
   Post_Max_Min = 681967,
   Y_Post_Max_Min = 2010,
   Y_RMean = 2014
 ) %>% 
   mutate(Status = ifelse(Y_RMean > Y_Post_Max_Min & Post_Max_Min < (Max_Peak*0.10) & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50)),"Rebuilding",
                          ifelse((Y_RMean < Y_Max_Peak & RMean <= (Max_Peak*0.50) | Y_Max_Peak == max(Y_RMean)),"Developing",
                                 ifelse(RMean > Max_Peak*0.50,"Max Exploited",
                                        ifelse(Y_RMean > Y_Max_Peak & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))," Over Exploited",
                                               ifelse(Y_RMean > Y_Max_Peak & RMean < Max_Peak*0.10,"Collapsed",
                                                      "NA")
                                        )
                                 )
                          )
   )
   )
